Practical Hilbert space approximate Bayesian Gaussian processes for probabilistic programming

### 2022-02-16 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
updates in section 4

@@

\begin{figure*}
\centering
\subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig6_7_relationships.png}}
\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ \hspace{0.1mm} ($c = \frac{L}{S}$) \hspace{0.1mm} and the length-scale normalized by the half-range of the data ($\frac{\ell}{S}$), for the square exponential, Mat\'ern ($\nu$=3/2) and Mat\'ern ($\nu$=5/2) covariance functions.}
  \label{fig6_7_relationships}
\end{figure*}

Figure \ref{fig6_7_relationships} depicts how these three factors interact and affect the accuracy of the HSGP approximation for a GP with square exponential covariance function, Mat\'ern ($\nu$=3/2) and Mat\'ern ($\nu$=5/2) covariance functions and a single input dimension. More precisely, for a given GP model (with a particular covariance function) with length-scale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_7_relationships} shows the minimum number of basis functions $m$ required to obtain an accurate approximation in the sense of satisfying eq. (\ref{eq_diff_covs}). We considered an approximation to be a sufficiently accurate when the total variation difference between the approximate and exact covariance functions, $\varepsilon$ in eq. (\ref{eq_diff_covs}), is below 1$\%$ of the total area under the curve of the exact covariance function $k$ such that
%
\begin{equation} \label{eq:rel_total_distance}
 \int | k(\tau) - \tilde{k}_m(\tau)|  \,\mathrm{d}\tau  < 0.01 \int k(\tau) \,\mathrm{d}\tau,
\end{equation}
where $\tilde{k}_m$ is the approximate covariance function with $m$ basis functions. Alternatively, these figures can be understood as providing the minimum $c$ that we should use for given $\ell$ and $m$. Of course, we may also read it as providing the minimum $\ell$ that can be approximated with high accuracy given $m$ and $c$. We obtain the following main conclusions:

\begin{itemize}
\item As $\ell$ increases, $m$ required for an accurate approximation decrease.
\item The lower $c$, the smaller $m$ can and $\ell$ must be to achieve an accurate approximation.
\item For a given $\ell$ there exist a minimum $c$ under which a accurate approximation is never going to be achieved regardless of $m$. This fact can be seen in Figures~\ref{fig6_7_relationships} and \ref{fig6_7_relationships} as the contour lines which represent $c$ have an end in function of $\ell$ (Valid $c$ are restricted in function of $\ell$). As $\ell$ increases, the minimum valid $c$ also increases.
\end{itemize}

\subsubsection{Numerical equations}
\label{num_equations}

From the empirical observations the hypothesis of linear proportionality between $m$, $\frac{l}{S}$ (lenghscale $l$ relative to the half range of the input domain $S$) and $c$ has been proved. Following we show the numerical functions governing these relationships for every kernel.

\paragraph*{Squared exponential}:
%
\begin{align}\label{eq_m_l_QE}
&m= 1.75 \, \frac{c}{\nicefrac{l}{S}} \;\; \Leftrightarrow \;\; \nicefrac{l}{S}= 1.75 \, \frac{c}{m},
\end{align}
%
with
%
\begin{align}\label{eq_c_vs_l_QE}
&c \geq \, 3.2 \, \nicefrac{l}{S} \;\; \& \;\, c \geq \, 1.2      
\end{align}

\paragraph*{Mat\'ern ($\nu$=3/2)}:
%
\begin{align}\label{eq_m_l_mat32}
&m= 3.42 \, \frac{c}{\nicefrac{l}{S}} \;\; \Leftrightarrow \;\; \nicefrac{l}{S}= 3.42 \, \frac{c}{m},
\end{align}
%
with
%
\begin{align}\label{eq_c_vs_l_mat32}
&c \, \geq \, 4.5 \, \nicefrac{l}{S} \;\; \& \;\, c \geq \, 1.2        
\end{align}

\paragraph*{Mat\'ern ($\nu$=5/2)}:
%
\begin{align}\label{eq_m_l_mat52}
&m= 2.65 \, \frac{c}{\nicefrac{l}{S}} \;\; \Leftrightarrow \;\; \nicefrac{l}{S}= 2.65 \, \frac{c}{m},
\end{align}
%
with
%
\begin{align}\label{eq_c_vs_l_mat52}
&c \, \geq \, 4.1 \, \nicefrac{l}{S} \;\; \& \;\, c \geq \, 1.2       
\end{align}

Notice that we establish a lower bound for $c$ set to 1.2.

Figure~\ref{fig6_7_relationships} was build for a GP with a unidimensional covariance function, which result in a surface depending on three variables, $m$, $c$ and $\ell$. An equivalent figure for a GP model with a two-dimensional covariance function would result in a surface depending on four variables, $m$, $c$, $\ell_1$ and $\ell_2$, which is more difficult to be graphically represented. More precisely, in the multi-dimensional case, whether the approximation is close enough might depend only on the ratio between wigglyness in every dimensions. For instance, in the two-dimensional case, it would depend on the ratio between $\ell_1$ and $\ell_2$ and could be graphically represented. Future research will focus on building useful graphs or analytical models that provide these relations in multi-dimensional cases. However, as an approximation, we can use the unidimensional GP conclusions in Figures \ref{fig6_7_relationships} and \ref{fig6_7_relationships} to check the accuracy by analyzing individually the different dimensions of a multidimensional GP model.

@@

As stated above, Figure~\ref{fig6_7_relationships} provides the minimum length-scale that can be closely approximated given $m$ and $c$. This information serves as a powerful diagnostic tool in determining if the obtained accuracy is acceptable. As the length-scale $\ell$ controls the wigglyness of the function, it strongly influences the difficulty of obtaining accurate inference about the function from the data. Basically, if the length-scale estimate is accurate, we can expect the HSGP approximation to be accurate as well.

Having obtained an estimate $\hat{\ell}$ of $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ exceeds the minimum length-scale provided in Figure~\ref{fig6_7_relationships} (depending on which kernel is used). If $\hat{\ell}$ exceeds this recommended minimum length-scale, the approximation is assumed to be good. If, however, $\hat{\ell}$ does not exceed recommended minimum length-scale, the approximation may be inaccurate and $m$ should be increased or $c$ decreased. We may also use this diagnostic in an iterative procedure by starting from some initial guess of $\ell$ and initial values for $m$ and $c$, and if the estimated $\hat{\ell}$ is below the minimum length-scale, repeat the process while increasing $m$ or decreasing $c$. As mentioned earlier, $c$ cannot be decreased too much as the lowest useful value of $c$ is restricted by the length-scale. Thus, increasing $m$ may usually the preferred approach.

If we look back to the conclusions drawn from Figures~\ref{fig4_MSE_vs_J} and \ref{fig5_lscale_vs_J}, where $m = 10$ basis functions and a boundary factor of $c = 1.5$ were sufficient to obtain an accurate approximation of a function with $\ell = 0.3$, we can recognize that these conclusions also matches those obtained from Figure~\ref{fig6_7_relationships}.

\subsubsection{A step-by-step user-guide to apply the diagnostics}

Based on the above proposed diagnostics, we obtain a simple step-by-step procedure that users can follow in order to obtain an accurate HSGP approximation. The procedure is split into two phases, Phase A and B, which have to be completed consequtively.
%In the following, we will refer to the Equations from Section XXX for the double-exponential kernel, but the same procedure works for other kernels by using the corresponding Equations provided in the same Section.

@@

\begin{enumerate}
\item[Step A1.] Make an initial guess on true the lengthscale $l_1$ of the function to be learned. We recommend a large lengthscale, about $l_1 \approx 1$. Set the iteration index to $i=1$.
%\item[] \hspace{-1cm} \textit{(Initial iteration, $i=1$)}
\item[Step A2.] Obtain the minimum valid boundary factor $c_i$ determined by $l = l_i$ for the given kernel as per Section \ref{num_equations} 
\item[Step A3.] Obtain the mimimum valid number of basis functions $m_i$ determined by $l = l_i$ and $c = c_i$ for the given kernel as per Section \ref{num_equations}. 
\item[Step A4.] Fit an HSGP model using $m_i$ and $c_i$ and ensure convergence of the MCMC chains.
\item[Step A5.] Perform the lengthscale diagnostic by checking if $\hat{l}_i \pm 0.02 \geq l_i$.

If the diagnostic is FALSE, the HSGP approximation is not yet sufficiently accurate. Set $l_{i+1} = \hat{l}_{i}$, increase the iteration index $i = i + 1$, and go back to Step A2.

@@

\item[Step B2.] Set $m_{i+1} = m_{i} + 10$ and increase the iteration index $i = i + 1$.

\item[Step B3.] Obtain the minimum valid boundary factor $c_i$ determined by $l = \hat{l}_{i-1}$ for the given kernel as per Section \ref{num_equations}.

\item[Step B4.] Obtain the mimimum valid lengthscale $l_i$ that can be accurately estimated by $m = m_i$ and $c = c_i$ for the given kernel as per Section \ref{num_equations}.

@@

\item[Step B6.] Perform the lengthscale diagnostic by checking if $\hat{l}_i \pm 0.02 \geq l_i$.

Check the stability of both $\hat{l}_i$ and the measures of predictive accuracy relative to the previous iteration.

@@

\begin{figure*}
\centering
\subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig_sec45_comparing_lscale.png}}
\caption{Estimated exact GP and approximate HSGP lengthscales for various datasets with varying smoothness.}
  \label{fig_sec45_comparing_lscale}
\end{figure*}

Figure \ref{fig_sec45_comparing_lscale}-left shows a comparison of the length-scale estimates
obtained from the exact GP and HSGP models with a squared exponential kernel from various noisy datasets drawn from underlying functions with varying smoothness. Different values of the number of basis functions $m$ are used when estimating the HSGP models, and the boundary factor $c$ is set to a valid and optimum value in every case by using equation (\ref{eq_c_vs_l_QE}). Figure \ref{fig_sec45_comparing_lscale}-right shows the RMSE of the HSGP models with the exact GP model as the reference.

Using large number of basis functions in smooth datasets (with length-scales greater than 0.5) causes overestimating the underlying functions slightly and so the length-scale too (Figure \ref{fig_sec45_comparing_lscale}-left). Here it comes out a possible weaknesses of the approach that when we want to fit very smooth functions we should try not to use too many basis functions than necessary because we will run into overestimating problems. Apart from that, an accurate estimate of the length-scale (Figure \ref{fig_sec45_comparing_lscale}-left) implies a small RMSE (Figure \ref{fig_sec45_comparing_lscale}-right).

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig_sec45_2_oversmoothing_effect.png}
\caption{Oversmooth posterior functions when using large number of basis functions for smooth datasets.}
  \label{fig_sec45_2_oversmoothing_effect}
\end{figure}

Next we apply the diagnosis procedure over all the datasets of Figure \ref{fig_sec45_comparing_lscale}-left and we show that following our recomendations and the diagnosis we get the optimum solution with minimum requirements.

>>>>

\subsection{Linear representation of a periodic squared exponential covariance function} \label{sec_method_periodic}

A GP model with a periodic covariance function does no fit in the framework of the HSGP approximation covered in this study as a periodic covariance function has not a spectral representation, but it has also a low-rank representation. In Appendix~\ref{sec_periodic}, we briefly present the approximate linear representation of a periodic squared exponential covariance function as developed by \citet{solin2014explicit}, analyze the accuracy of this approximation and, finally, derive the GP model with this approximate periodic square exponential covariance function.

@@

\begin{figure*}
\centering
\subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig6_relationships.png}}
\caption{Relation among the minimum number of basis functions $m$, the boundary factor $c$ \hspace{0.1mm} ($c = \frac{L}{S}$) \hspace{0.1mm} and the length-scale normalized by the half-range of the data ($\frac{\ell}{S}$), for the square exponential, Mat\'ern ($\nu$=3/2) and Mat\'ern ($\nu$=5/2) covariance functions.}
  \label{fig6_relationships}
\end{figure*}

Empirical simulations are carried out to analyze the relationships between $m$, $c$ and $l$. Figure \ref{fig6_relationships} depicts how these three factors interact and affect the accuracy of the HSGP approximation for a GP with square exponential, Mat\'ern ($\nu$=3/2) and Mat\'ern ($\nu$=5/2) covariance functions and a single input dimension. More precisely, for a given GP model (with a particular covariance function) with length-scale $\ell$ and given a boundary factor $c$, Figure \ref{fig6_relationships} shows the minimum number of basis functions $m$ required to obtain an accurate approximation in the sense of satisfying eq. (\ref{eq_diff_covs}). We considered an approximation to be a sufficiently accurate when the total variation difference between the approximate and exact covariance functions, $\varepsilon$ in eq. (\ref{eq_diff_covs}), is below 1$\%$ of the total area under the curve of the exact covariance function $k$ such that
%
\begin{equation} \label{eq:rel_total_distance}
 \int | k(\tau) - \tilde{k}_m(\tau)|  \,\mathrm{d}\tau  < 0.01 \int k(\tau) \,\mathrm{d}\tau,
\end{equation}
where $\tilde{k}_m$ is the approximate covariance function with $m$ basis functions. Alternatively, these figures can be understood as providing the minimum $c$ that we should use for given $\ell$ and $m$. Of course, we may also read it as providing the minimum $\ell$ that can be approximated with high accuracy given $m$ and $c$. We obtain the following main conclusions:

\begin{itemize}
\item As $\ell$ increases, $m$ required for an accurate approximation decrease.
\item The lower $c$, the smaller $m$ can and $\ell$ must be to achieve an accurate approximation.
\item For a given $\ell$ there exist a minimum $c$ under which a accurate approximation is never going to be achieved regardless of $m$. This fact can be seen in Figure~\ref{fig6_relationships} as the contour lines which represent $c$ have an end in function of $\ell$ (Valid $c$ are restricted in function of $\ell$). As $\ell$ increases, the minimum valid $c$ also increases.
\end{itemize}

\subsubsection{Relationships between $J$ and $\ell$ for a periodic squared exponential covariance function} \label{subsec_empiric_relations_periodic}

In Appendix~\ref{sec_periodic}, we analyze the accuracy of the linear representation of a periodic squared exponential covariance function and derive the minimum number of terms $J$ in the approximation required to achieve a close approximation to the exact periodic squared exponential kernel as a function of the length-scale $\ell$ of the kernel. Since this is a series expansion of sinusoidal functions, the approximation does not depend on any boundary condition. This relationship between $J$ and $\ell$ is gathered in Figure~\ref{figB1_m_lscale_periodic}.

\subsubsection{Numerical equations}
\label{sec_num_equations}

From the empirical observations, numerical equations governing the relationships between $m$, $c$ and $\ell$ were estimated, which show an agreement with the hypothesis of linear proportionality between $m$, $\frac{l}{S}$ (lenghscale $l$ relative to the half range of the input domain $S$) and $c$. The estimated numerical equations for each particular covariance function are the following:

\paragraph*{Squared exponential}:
%
\begin{align}\label{eq_m_l_QE}
&m \geq 1.75 \, \frac{c}{\nicefrac{l}{S}} \;\; \Leftrightarrow \;\; \nicefrac{l}{S} \geq 1.75 \, \frac{c}{m},
\end{align}
%
with
%
\begin{align}\label{eq_c_vs_l_QE}
&c \geq 3.2 \, \nicefrac{l}{S} \;\; \& \;\, c \geq 1.2      
\end{align}

\paragraph*{Mat\'ern ($\nu$=3/2)}:
%
\begin{align}\label{eq_m_l_mat32}
&m \geq 3.42 \, \frac{c}{\nicefrac{l}{S}} \;\; \Leftrightarrow \;\; \nicefrac{l}{S} \geq 3.42 \, \frac{c}{m},
\end{align}
%
with
%
\begin{align}\label{eq_c_vs_l_mat32}
&c \geq 4.5 \, \nicefrac{l}{S} \;\; \& \;\, c \geq 1.2        
\end{align}

\paragraph*{Mat\'ern ($\nu$=5/2)}:
%
\begin{align}\label{eq_m_l_mat52}
&m \geq 2.65 \, \frac{c}{\nicefrac{l}{S}} \;\; \Leftrightarrow \;\; \nicefrac{l}{S} \geq 2.65 \, \frac{c}{m},
\end{align}
%
with
%
\begin{align}\label{eq_c_vs_l_mat52}
&c \geq 4.1 \, \nicefrac{l}{S} \;\; \& \;\, c \geq 1.2       
\end{align}

Notice that we establish a lower bound at 1.2 for $c$ in every case.

Figure~\ref{fig6_relationships} and previous equations (\ref{eq_m_l_QE})-(\ref{eq_c_vs_l_mat52}) were obtained for a GP with a unidimensional covariance function, which result in a surfaces depending on three variables, $m$, $c$ and $\ell$. Equivalent results for a GP model with a two-dimensional covariance function would result in a surface depending on four variables, $m$, $c$, $\ell_1$ and $\ell_2$. More precisely, in the multi-dimensional case, whether the approximation is close enough might depend only on the ratio between wigglyness in every dimensions. For instance, in the two-dimensional case, it would depend on the ratio between $\ell_1$ and $\ell_2$. Future research will focus on building useful graphs or analytical models that provide these relations in multi-dimensional cases. However, as an approximation, we can use the unidimensional GP conclusions in Figure \ref{fig6_relationships} or equations (\ref{eq_m_l_QE})-(\ref{eq_c_vs_l_mat52}) to check the accuracy by analyzing individually the different dimensions of a multidimensional GP model.

\subsubsection{Numerical equation for a periodic squared exponential covariance function} \label{sec_num_equation_periodic}

The numerical equation governing the relationships between $J$ and $\ell$ for a periodic squared exponential covariance function was estimated in eq.~\ref{eq_J_l_periodic} which is depicted next:
%
\begin{align*}
&J \geq \frac{3.72}{\ell} \;\; \Leftrightarrow \;\; \ell \geq \frac{3.72}{J}
\end{align*}

@@

As stated above, equations (\ref{eq_m_l_QE}), (\ref{eq_m_l_mat32}) and (\ref{eq_m_l_mat52}) (depending on which kernel is used) in previous Section~\ref{sec_num_equations} provide the minimum length-scale that can be closely approximated given $m$ and $c$. This information serves as a powerful diagnostic tool in determining if the obtained accuracy is acceptable. As the length-scale $\ell$ controls the wigglyness of the function, it strongly influences the difficulty of obtaining accurate inference about the function from the data. Basically, if the length-scale estimate is accurate, we can expect the HSGP approximation to be accurate as well.

Having obtained an estimate $\hat{\ell}$ of $\ell$ from the HSGP model based on prespecified $m$ and $c$, we can check whether or not $\hat{\ell}$ exceeds the minimum length-scale provided by equations (\ref{eq_m_l_QE}), (\ref{eq_m_l_mat32}) and (\ref{eq_m_l_mat52}) (depending on which kernel is used). If $\hat{\ell}$ exceeds this recommended minimum length-scale, the approximation is assumed to be good. If, however, $\hat{\ell}$ does not exceed recommended minimum length-scale, the approximation may be inaccurate and $m$ and/or $c$ need to be updated. If we look back to the conclusions drawn from Figures~\ref{fig4_MSE_vs_J} and \ref{fig5_lscale_vs_J}, where $m = 10$ basis functions and a boundary factor of $c = 1.5$ were sufficient to obtain an accurate approximation of a function with $\ell = 0.3$, we can recognize that these conclusions also matches those obtained from eq. (\ref{eq_m_l_QE}).

Apart from providing a powerful diagnostic tool in determining if the approxiamtion is sufficiently accurate, equations in previous Section~\ref{sec_num_equations} also give the optimal values for $m$ (the minimum $m$ required for an accurate approximation) and $c$ (the minimum $c$ that allows for the minimum $m$) that allow to spend minimum computations.

Guidelines for updating $m$ and/or $c$ must be provided:

\begin{itemize}
	\item $c$ must be big enough for the length-scale $l$ of the function to be learned and,
	\item $m$ must be big enough as a function of $\ell$ and $c$. 
\end{itemize}

We recommend using this diagnostic in an iterative procedure by starting from some initial guess of $\ell$ and obtain the optimum valid values for $c$ and $m$ as a function of this initial guess of $\ell$ by using equations per Section~\ref{sec_num_equations} , and if the estimated $\hat{\ell}$ is below the minimum length-scale determined by $c$ and $m$ provided by equations in Section~\ref{sec_num_equations}, update $\ell$ with $\hat{\ell}$ and repeat the process.

A detailed guideline with the steps to perform diagnosis, ensuring that a good approximation is achieved with the minimum computational requirements (i.e. minimum number of basis functions $m$) is provided in  next Section~\ref{sec_user_guide}.

\subsubsection{A step-by-step user-guide to apply the diagnostics} \label{sec_user_guide}

Based on the above proposed diagnostics, we obtain a simple step-by-step procedure that users can follow in order to obtain an accurate HSGP approximation. The procedure is split into two phases, Phase A and B, which have to be completed consecutively.
%In the following, we will refer to the Equations from Section XXX for the double-exponential kernel, but the same procedure works for other kernels by using the corresponding Equations provided in the same Section.

@@

\item[Step A1.] Make an initial guess on the true length-scale $\ell_1$ of the function to be learned. If there is no approximate prior information, we recommend a medium-large length-scale, about $\ell_1 \approx [0.5-1]$. Set the iteration index to $i=1$.

\item[Step A2.] Obtain the minimum valid boundary factor $c_i$ determined by $\ell = \ell_i$ for the given kernel as per Section \ref{sec_num_equations}

\item[Step A3.] Obtain the mimimum valid number of basis functions $m_i$ determined by $\ell = \ell_i$ and $c = c_i$ for the given kernel as per Section \ref{sec_num_equations}.

\item[Step A4.] Fit an HSGP model using $m_i$ and $c_i$ and ensure convergence of the MCMC chains.

\item[Step A5.] Perform the lengthscale diagnostic by checking if $\hat{\ell}_i + 0.01 \geq \ell_i$.

If the diagnostic is FALSE, the HSGP approximation is not yet sufficiently accurate. Set $\ell_{i+1} = \hat{\ell}_{i}$, increase the iteration index $i = i + 1$, and go back to Step A2.

@@

\item[Step B2.] Set $m_{i+1} = m_{i} + 5$ and increase the iteration index $i = i + 1$.

\item[Step B3.] Obtain the minimum valid boundary factor $c_i$ determined by $\ell = \hat{\ell}_{i-1}$ for the given kernel as per Section \ref{sec_num_equations}.

\item[Step B4.] Obtain the mimimum valid lengthscale $\ell_i$ that can be accurately estimated by $m = m_i$ and $c = c_i$ for the given kernel as per Section \ref{sec_num_equations}.

@@

\item[Step B6.] Perform the lengthscale diagnostic by checking if $\hat{\ell}_i + 0.01 \geq \ell_i$.

Check the stability of both $\hat{\ell}_i$ and the measures of predictive accuracy relative to the previous iteration.

@@

\begin{figure*}
\centering
\subfigure{\includegraphics[width=\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig7_lscale_comparison.png}}
\caption{Exact GP length-scale ($\ell_{GP}$) and approximate HSGP lengthscale ($\hat{\ell}$) for various datasets with varying smoothness (left) and their corresponding root mean squared errors (RMSE) computed with the exact GP model as the reference.}
  \label{fig7_lscale_comparison}
\end{figure*}

First of all in this section we illustrate that accurate estimates of the length-scale implies accurate approximations with a small RMSEs. Thus, Figure~\ref{fig7_lscale_comparison}-left shows a comparison of the length-scale estimates
obtained from the exact GP and HSGP models with a squared exponential kernel, from various noisy datasets drawn from underlying functions with varying smoothness. Different values for the number of basis functions $m$ are used when estimating the HSGP models, and the boundary factor $c$ is set to a valid and optimum value in every case by using eq. (\ref{eq_c_vs_l_QE}). Figure \ref{fig7_lscale_comparison}-right shows the RMSE of the HSGP models with the exact GP model as the reference. It can be seen that accurate estimates of the length-scale implies small RMSEs.

Furthermore, from Figure~\ref{fig7_lscale_comparison}-left it can be seen that using large number of basis functions relative to the wigglyness of the function to be learned (i.e. true length-scale) causes overestimating the underlying functions (see Figure~\ref{fig8_oversmoothed_posterios} as an illustration of the oversmoothing effect as a function of $m$ in a moderate smooth dataset) and, consequently, on the length-scale estimate too (Figure \ref{fig7_lscale_comparison}-left). Here it comes out a possible weaknesses of the approach which consists in using too many basis functions relative to the true length-scale causes overestimating the posteriors.

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth, trim = 0mm 0mm 0mm 0mm, clip]{fig8_oversmoothed_posterios.png}
\caption{Oversmoothing effect on posterior functions when using too many number of basis functions relative to the wigglyness of the function to be learned.}
  \label{fig8_oversmoothed_posterios}
\end{figure}

Secondly in this section, Figure~\ref{fig9_diagnosis} shows the iterative steps of applying the diagnosis procedure explained in Section~\ref{sec_user_guide} over some of the data sets also used in the analysis in Figure~\ref{fig7_lscale_comparison}-left, showing that by following our recommendations, an optimum solution with minimum computational requirements is achieved. Figure~\ref{fig10_diagnosis_lscale_comparison} graphically compares the exact GP length-scale and the estimated HSGP length-scale in every iteration and data set. Between 2 and 4 iterations, depending on wigglyness of the function to be learned and the distance between the initial guess of the length scale and the true length scale, are sufficient to reach the optimal values of $m$ and $c$. Furthermore, as can be seen in Figure~\ref{fig10_diagnosis_lscale_comparison}, following our diagnostic recommendations avoids overestimation.

As examples, the iterative steps applied to perform diagnosis on two of the data sets in Figure~\ref{fig9_diagnosis} are described next:

\paragraph*{Data set with the exact GP length-scale $\ell_{\text{GP}} = 0.08$:} 
\begin{itemize}
\item[] Iteration 1:
	\begin{enumerate}
	\item Making the initial guess that the length-scale $\ell_1=0.5$.

	\item Computing the optimal $c_1$ as a function of $\ell_1$ by using eq.~(\ref{eq_c_vs_l_QE}).

	\item Computing the minimum $m_1$ as a function of $c_1$ and $\ell_1$ by using eq.~(\ref{eq_m_l_QE}).

	\item Fitting the HSGP model and obtain the estimated $\hat{\ell}_1$.

	\item The diagnostic $\hat{\ell}_1 + 0.01 \geq \ell_1$ gives FALSE.
	\end{enumerate}

\item[] Iteration 2:
	\begin{enumerate}
	\item As the diagnostic in previous iteration 1 was FALSE, update $\ell_2$ with $\hat{\ell}_1$.

	\item Computing the optimal $c_2$ as a function of $\ell_2$ and $m_1$ as a function of $c_2$ and $\ell_2$ by using equations~(\ref{eq_c_vs_l_QE}) and eq.~(\ref{eq_m_l_QE}), respectively.

	\item Fitting the HSGP model using $c_2$ and $m_2$ and estimating $\hat{\ell}_2$.

	\item The diagnostic $\hat{\ell}_2 + 0.01 \geq \ell_2$ gives FALSE.
	\end{enumerate}

\item[] Iteration 3:
	\begin{enumerate}
	\item As the diagnostic in previous iteration 2 was FALSE, updating $\ell_3$ with $\hat{\ell}_2$ and repeat the process as in iteration 2.

	\item The diagnostic $\hat{\ell}_3 + 0.01 \geq \ell_3$ gives TRUE.
	\end{enumerate}

\item[] Iteration 4:
	\begin{enumerate}
	\item As the diagnostic in step 4 in previous iteration 3 was TRUE, increasing $m_4 = m_3 + 5$.

	\item Computing the optimal $c_4$ as a function of $\ell_3$ by using eq.~(\ref{eq_c_vs_l_QE}).

	\item Fitting the HSGP model using $c_4$ and $m_4$ and estimating $\hat{\ell}_4$.

	\item The diagnostic $\hat{\ell}_4 + 0.01 \geq \ell_4$ gives TRUE.

	\item As RMSE, $R^2$ and ELPD are stable relative to previous iteration 4 the approximation must be sufficiently accurate.
	\end{enumerate}

\end{itemize}

\paragraph*{Data set with the exact GP length-scale $\ell_{\text{GP}} = 1.4$:} 
\begin{itemize}
\item[] Iteration 1:
	\begin{enumerate}
	\item Making the initial guess that the length-scale $\ell_1=1$.

	\item Computing the optimal $c_1$ as a function of $\ell_1$ by using eq.~(\ref{eq_c_vs_l_QE}).

	\item Computing the minimum $m_1$ as a function of $c_1$ and $\ell_1$ by using eq.~(\ref{eq_m_l_QE}).

	\item Fitting the HSGP model and obtain the estimated $\hat{\ell}_1$.

	\item The diagnostic $\hat{\ell}_1 + 0.01 \geq \ell_1$ gives TRUE.
	\end{enumerate}

\item[] Iteration 2:
	\begin{enumerate}
	\item As the diagnostic in previous iteration 1 was TRUE, increasing $m_2 = m_1 + 5$.

	\item Computing the optimal $c_2$ as a function of $\ell_2$ by using eq.~(\ref{eq_c_vs_l_QE}).

	\item Fitting the HSGP model using $c_2$ and $m_2$ and estimating $\hat{\ell}_2$.

	\item The diagnostic $\hat{\ell}_2 + 0.01 \geq \ell_2$ gives TRUE.

	\item As RMSE, $R^2$ and ELPD are stable relative to previous iteration 2 the approximation must be sufficiently accurate.
	\end{enumerate}

\end{itemize}

It is worth emphasizing that using the optimum values for $m$ and $c$ that allow for minimum computational requirements can be essential in uni-dimensional cases with very large data sets, and even more in low-dimensional cases ($D \leq 4$; Notice from section~\ref{sec_method_multi} that for $D > 4$ it starts to be unpractical in any case) even with small data sets, since knowing the optimum values for $m$ and $c$ for each individual dimension can reduce significantly the computational requirements in these multi-dimensional cases, otherwise they would be unfeasible to be fitted.
